---
title: "R_Assignment_Group 5"
Members: Ananda Chatterjee 
         Avinash Yadav 
         Dhruv Sobti 
         Hrisav Bhowmick 
         Vineet Kumar
---

=============================================================================================================================================================================================
Given:
      - A IVF Hospital Chain has 1000 patients/month capacity with 100 patients/month per Center so hospital has 10 Centers
      - Each Center patient footfall follows uniform distribution with footfalls lying between 72 to 84 per center
      - Hospital makes a profit of $1000 for each baby born, $200 for a twin, $-1000 for a triplet, $-2000 for quadruplet and 0 profit when there is a miscarriage
      - Probability of baby being born is 65%, for twin - 25%, triplet - 5%, Quadruplet - 1% and for miscarriage it is 4%

Find : Detailed projection of Profits made by IVF chain , Analysis of footfalls , annual Budget(Individual Assignment)

=============================================================================================================================================================================================
=============================================================================================================================================================================================
PART 1 - GROUP ASSIGNMENT  :  Profit Projection of Month


Solution:


Function : with 3 parameters 1st is required , other 2 have default values 
           Calculating Distribution of Birth, Number of born Babies, Twins, Triplets, Quadruplets and Miscarriages of each IVF Hospital Center.
           This Function is also calculating Each Center Profit per birth and also total profit from every Center
```{r}

#function with default values for probability and payoff
center = function(footfall,probOfBirth=c(0.65,0.25,0.05,0.01,0.04),payoff=c(1000,200,-1000,-2000,0))
{
  outcome=c('baby','twin','triplet','quadruplet','miscarriage')                 #different Outcome that each Patient can possibly have
  
  #Creating a Distribution of Birth outcome by sampling via the 'outcome', with size = average patient footfalls and probability of birth rates respectively
  distPerCenter=sample(outcome,footfall,replace = TRUE,prob = probOfBirth)
  
  babyCount=sum(distPerCenter=='baby')                                          #Counting no. of babies born
  twinCount=sum(distPerCenter=='twin')                                          #Counting no. of Twins Born
  tripletCount=sum(distPerCenter=='triplet')                                    #Counting no. of triplets Born
  quadrupletCount=sum(distPerCenter=='quadruplet')                              #Counting no. of quadruplets Born
  miscarriageCount=sum(distPerCenter=='miscarriage')                            #Counting no. of Miscarriages happened
  
  #Calculating the Profit or Loss made from each Outcome by multiplying the count of outcomes with respective payoffs
  profitPerOutcome=c(babyCount,twinCount,tripletCount,quadrupletCount,miscarriageCount)*payoff       
  profitPerCenter=sum(profitPerOutcome)                                         #Calculating Total Profit made From one center                           
  
  #Returning list of Outcome frequency(starting 5 values),Patients coming at the particular center(6th value),Profit made by the Center(7th value) and Distribution of birth outcome(8th value) 
  return ( list(babyCount,twinCount,tripletCount,quadrupletCount,miscarriageCount,footfall,profitPerCenter,distPerCenter) )
}
```

Finding Footfalls of every Center , then calling the center() function to find the baby Outcome for each footfall for every Center

```{r}

distAllCenter=list()                                                            #list to store Distribution From all Centers

cnames = c("Baby","Twin","Triplet","Quadruplet","Miscarriage","Footfall","Total Profit")
rnames = c(1:10)
rnames[1:10] = paste("Center -",rnames[1:10])
#Matrix to store records of All Center with 10 rows for Center and 7 Col. for each outcome(5),footfall(6th),Profit(7th)
IVFMatrix = matrix(0,byrow=TRUE,nrow=10,ncol=7, dimnames=list(rnames,cnames))      

#Looping 10 times for each center
for (centrNo in seq(1:10))
{
  footfall=round(runif(1,72,84))                                                #Calculating footfall for each center according to given Condition (uniform Distribution between(72-84))
  centerOutcome=center(footfall)                                                #Calling center() and storing all 8 values returned from the function in the variable 
  distAllCenter[[paste('Center',centrNo)]]=centerOutcome[[8]]                   #Storing distribution of birth outcome of each center with tag - 'Center 1','Center 2'...'Center 10'
  
  #looping for first 7 seven values in the returned list from center()
  for (value in c(1:7))
  {
   IVFMatrix[centrNo,value]=centerOutcome[[value]]                              #filling the Matrix with corresponding returned values from center() 
  }
  
  #Printing Footfall for each center and its Distribution of Births
  print(paste('Footfall of Center ',centrNo,':',centerOutcome[[6]]))  
  print(paste('Outcome of Center ',centrNo,':'))
  print(distAllCenter[[paste('Center',centrNo)]])
  cat('\n\n')
  
}

```


```{r}
#Records Of Each Center

cat("Records of Every Center in a Month\n\n")
IVFMatrix
totalProfit=sum(IVFMatrix[,7])                                                  #Calculating Total Profit For a month by summing all values from 7th Column in the records Matrix
meanProfit=mean(IVFMatrix[,7])                                                  #Mean Profit of a Month   
sdProfit=round(sd(IVFMatrix[,7]))
print(paste(cat("\n\n"),'Total Profit of IVF Hospital in a month :',totalProfit,'With average profit in a month :',meanProfit,'with Standard Deviation :',sdProfit))

```




===============================================================================================================================================================================================
===============================================================================================================================================================================================
===============================================================================================================================================================================================
PART 2 - INDIVIDUAL ASSIGNMENT   :  Profit Projection for whole year

                                        - For even numbered centers , no Miscarriages, triplet rate increases to 9%
                                        - Analysis for each month and Detailed Profit Projection of each Month

```{r}
distmonth = list()   #distribution of all birth outcomes of all months stored in this list
vars = list()        #birth Outcomes, footfalls, profit from each center and month stored in this list
mprofit = c()       #profit from all months is stored in this vector
cnames = c("Baby","Twin","Triplet","Quadruplet","Miscarriage","Footfall","Total Profit")
rnames = c(1:10)
rnames[1:10] = paste("Center -",rnames[1:10])
monthly_Matrix = matrix(0,byrow=TRUE,nrow=10,ncol=7, dimnames=list(rnames,cnames))        #record of each Month is stored in matrix via loop
prob_evenCenters=c(0.65,0.25,0.09,0.01,0)     #birth probability for even centers
prob_oddCenters=c(0.65,0.25,0.05,0.01,0.04)   #birth probability for odd centers


for (month in c(1:12))  #looping through each month
{
  distAllCenter=list()     #list to store immediate distribution of every center
  for (centrNo in seq(1:10))
  {
    footfall=round(runif(1,72,84))  #this will give a footfall value to each center
    
    if(centrNo %% 2==0){
      centerOutcome=center(footfall,prob_evenCenters)  #store values for even center   
    }
    else{
      centerOutcome=center(footfall,prob_oddCenters)   #store values for odd center 
    }
    
    distAllCenter[[paste('Center',centrNo)]]=centerOutcome[[8]]  #stores distribution of immediate center from returned value

    for (value in c(1:7))     #looping through starting 7 values of returned values
    {
     monthly_Matrix[centrNo,value]=centerOutcome[[value]]   #matrix of returned values
    }
  }
  
  distmonth[[paste('Month',month)]] = distAllCenter[]   #merging all distribution of every center by month in one list 
  vars[[paste('Month',month)]] = monthly_Matrix   #has all records of center by month in one list 
  mprofit = append(mprofit,sum(monthly_Matrix[,7]))   #vector has profit of all centers by month
}

```


```{r}

for (month in c(1:12)) #looping through months for output
{
  print(paste('Records of Month',month,':'))
  print(vars[[paste('Month',month)]])
  print(paste('Profit of Month',month,':',mprofit[month]))
  cat('\n\n')
}

```

```{r}

profy=sum(mprofit)           #Total Profit in a Year
meanpm=mean(mprofit)     #Mean profit in a Year
sd=round(sd(mprofit))  #Standard Deviation of Profit in a year

print(paste('Yearly Profit the Hospital Chain made:',profy))
print(paste('Average Profit per month:',round(meanpm, digits=2)))
print(paste('Standard Deviation:',sd))

```
