
The classical problem
=====================================
A hare & tortoise went for a race of 200 kilometers. The hare is fast wants to win at any condition. He starts the journey at 12 km per hour. Also due to the extra energy he expends, he ends up reducing the speed by 2 km every hour. When the speed is at the verge of becoming 0, the hare sleeps for anything between 1 and 3 hours. When he wakes up again, regains the speed at 12 km per hour again and keeps losing at 2 km every hour like before before he goes for sleep. The cycle continues.
The tortoise starts the race at 2km per hour and steadily increases the pace by 5% per hour. Who wins the race?


Calculate the tortoise case
================================================
```{r}
# Initially setting the number of hours to 0
ht_Tortoise = 0

# Tortoise starts at a speed of 2 km / hour
rt_Tortoise = 2

dt_Tortoise = 0 # Distance initialized to 0

# Tortoise goes till he reaches 200 km
while(dt_Tortoise<200)
{
  # Adding the hourly rate to the already traversed distance
  dt_Tortoise = dt_Tortoise + rt_Tortoise
  
  # Increasing the tortoise rate by 5% every hour
  rt_Tortoise = rt_Tortoise * 1.05
  
  # Increment the hour
  ht_Tortoise = ht_Tortoise + 1
  
}

print(paste("The tortoise took",ht_Tortoise,"hours",sep=" "))
```

Calculate the hare case 
================================================
```{r}

# Initially setting the number of hours to 0
ht_Hare = 0

# Setting the initial speed to 12 km/hour
rt_Hare = 12

# Initialize the distance covered and hours taken by the hare to 0
dt_Hare = 0

# The rabbit covers distance of 200 km
while(dt_Hare<200)
{
  # Adding the hourly rate to the already traversed distance
  dt_Hare = dt_Hare + rt_Hare
  
  rt_Hare = rt_Hare - 2
  
  if (rt_Hare<=0)
  {
    # The hare goes for sleep between 1 and 3 hours
    ht_Hare = ht_Hare + sample(1:3,size=1,replace = TRUE)
    
    #Reinitialize the speed of the Hare
    rt_Hare = 12
  }
  
  # Increment the hour
  ht_Hare = ht_Hare + 1
  
}
  
print(paste("The Rabbit took",ht_Hare,"hours",sep=" "))

```

Finding who won
================================================
```{r}

if(ht_Hare<ht_Tortoise)
  print("The Rabbit won") else print("The Tortoise won")

```


Problem - The Drunk Casino Owner
========================================================
You are celebrating a successful summer internship at a Casino with 5 friends. Incidentally it was the owner's birthday and he came to your table and offered drinks. You had 10 shots each with the owner merrily before he belted out an offer.

" I like you guys - let's play a game to make this evening more memorable. Each of the shots you took costs $10. Select any 3 numbers of your choice between 6 and 36 and each of you Roll the die taking turns. If any of your number comes up, I agree to pay you $1000 for the throw. If otherwise, you pay me $20 as the price of the shot. Think about it and I will send Nina for confirmation"

While you were discussing possibilities, the beautiful Nina came to your table. You became more confused when with a mysterious smile she said " I suggest you test your luck on a 34, 26 and 36 ". Enchanted over her charm, you all readily agreed to the offer and Nina walked away.

Given you are the only one trained in Data Analytics in the team - deduce if Nina was playing fair?
(Inspired from popular Casino stories)

Problem - The Drunk Casino Owner
========================================================
While you were discussing possibilities, the beautiful Nina came to your table. You became more confused when with a mysterious smile she said " I suggest you test your luck on a 34, 26 and 36 ". Enchanted over her charm, you all readily agreed to the offer and Nina walked away.

Given you are the only one trained in Data Analytics in the team - deduce if Nina was playing fair?
(Inspired from popular Casino stories)

What business impact the problem has?
========================================================
- You stand a chance to lose money
- There is also a window to make some money
- Assess your chances of Win vs Lose

What should be your stance
========================================================
- Evaluate the cases where you stand to lose
- Evaluate the probability of the losses
- Calculate the expected value

Define the Logic you should use
========================================================
- Create the total possible outcome
- Find the number of cases favouring winning
- Find the number of cases favouring losing
- Calculate the respective probabilities
- Get the expected values

Identify the Data container
=======================================================
- A container to store the possible values from the throw of a die
- A container to store the value generated from the throw of 1 die
- A container to store the value generated from the throw of 6 dice
- A container to store the value generated from the 10 experiments
- A container to count the number of successful case
- A container to find the probabilty of winning and probability of losing

Identify few Procedures to use
=======================================================
- Procedure to randomly get a throw from a die
- Procedure to sum the throw for all the 6 dice
- Procedure to get the probabilities for the dice throw


Problem Prototype - Data Definition
=========================================================
- Let's build the Solution with only 2 dice
```{r}
# Defining the 2 Dice
Die1 = c(1:6)
Die2 = seq(1,6,1)
```

Problem Prototype - Procedures
=========================================================
- Let's simulate the procedure of throwing all the Dice
```{r}
# Throwing the 2 Dice
Throw1 = sample(Die1, size=1, replace=TRUE)
Throw2 = sample(Die2, size=1, replace = TRUE)
```

Introducing a new Function - Replicate
========================================================
- The "replicate" function in R runs an operation as specified
- You need to provide the number of times its desired to run along with the function
- The output of the function is generally stored in a vector
- The syntax is as below
- replicate(Number of Times to Run, Name of the Procedure with parameters)

Problem Prototype - Procedures
=========================================================
- Let's simulate the procedure of repeating Dice throws
```{r}
# Throwing the 2 Dice
Throw1_10 = replicate(10,sample(Die1, size=1, replace=TRUE))
Throw2_10 = replicate(10,sample(Die2, size=1, replace = TRUE))

# Showing the Output
print(Throw1_10)
print(Throw2_10)
```

Problem Prototype - Procedures
=========================================================
- Getting the sum of the Throws
```{r}
# Throwing the 2 Dice
Throw_Sum = Throw1_10 + Throw2_10
print(Throw_Sum)
```

Getting the number of events from a Sample Space
=========================================================
- For 1 die, the total sample space is 6
- For 2 dice, the total sample space is 36
- For 3 dice, the total sample space is 216
- For 4 dice, the total sample space is 1296


Problem Prototype - Calculate Proportions
=========================================================
- Getting the Probabilities for 6 & 7
```{r}
# Getting the outputs which are 6 or 7
OP6or7 = (Throw_Sum==6|Throw_Sum==7)
print(OP6or7)

# Counting the number of cases for 6 or 7
Countof6or7 = sum(OP6or7)
print(Countof6or7)

WinProp = Countof6or7/length(Throw_Sum)
LoseProp = 1 - WinProp
```

Why we used the function "which" here
=====================================================================
- The "which" function gives you the index from a vector where matches
- You need to provide the vector name with the corresponding logical condition
- The output of the function is stored in a vector with only index of matching condition
- The syntax is as below
- which(Condition)

Calculating the Number of Cases from the Sample Space for the Prototype
====================================================================
- Do a Random Experiment
```{r}
# Getting the outputs which are 6 or 7
Throw1 = sample(Die1, size=600, replace=TRUE)
Throw2 = sample(Die2, size=600, replace = TRUE)
Throw_Sum = Throw1 + Throw2
OP6or7 = which(Throw_Sum==6|Throw_Sum==7)

# Counting the number of cases for 6 or 7
Throw6or7 = length(OP6or7)
```

Calculating the Probability from the Sample Space for the Prototype
====================================================================
- Use the Counts
```{r}
# Finding the Winning Probability when the Outputs are 6 or 7
WinProb = Throw6or7/length(Throw_Sum)
LoseProb = 1 - WinProb
print(WinProb)

#Theoretical Win Probability is as follows:
Casesfor6 = 5
Casesfor7 = 6
ActualProb = (Casesfor7+Casesfor6)/36
print(ActualProb)
```

Introducing to writing your own Function
========================================================
- Define the name of the function
- Tell the system that you are defining a function
- Syntax is [Function_Name = function()]
- If there are parameters that the function requires, define them within the bracket in function(parameters1, Param2,...)
- The block of Function starts with a curly bracket "{" and ends with a curly bracket "}"
- If the function returns something, mention that explicitly with return
- Syntax return(variable)
- Close the loop
- When calling the function, give the function name and any parameter that you may need



Problem Prototype - Writing a Function
=========================================================
- A function to simulate the Dice throwing

```{r}
# Declare the Function
Dice_Throw_Ind0 = function (Throw_Num_times)
{
  pDie = c(1:6)
  
  #Throw the Dice for the number of times you specify
  Throw_Dice = sample(pDie,size = Throw_Num_times,replace = TRUE)
    
  return(Throw_Dice)
  
}

print(Dice_Throw_Ind0(10))
  
```

Problem Prototype - Writing a Function
=========================================================
- A function to simulate the Dice throwing

```{r}
# Declare the Function
Dice_Throw_Ind = function (Throw_Num_times)
{
  pDie = c(1:6)
  
  #Throw the Dice once and replicate the opeartion for the number of times you specify
  Throw_Dice = replicate(Throw_Num_times, sample(pDie,size=1,replace = TRUE))
    
  return(Throw_Dice)
  
}

print(Dice_Throw_Ind(10))
  
```



Problem Prototype - Calling a Function from a Function
=========================================================
- A function to simulate the summing of the Dice

```{r}
# Declare the Function
Dice_Throw_Grp = function (Throw_Num,Dice_Num) {
  counter = 0
  Throw_Sum = rep(0,Throw_Num)
  
  while (counter < Dice_Num){
    
    #Throw the Dice for the number of times you specify
    Throw_Dice = Dice_Throw_Ind(Throw_Num_times = Throw_Num )
    
    #Sum up the Throws
    Throw_Sum = Throw_Sum + Throw_Dice
    
    #Add the Counter
    counter = counter + 1
  }
  return(Throw_Sum)
}
```


Problem Prototype - Writing a Complex Function
=========================================================
- A function to simulate the summing of the Dice
```{r}
# Declare the Function
Dice_Throw_Sum = function (Throw_Num,Dice_Num) {
  counter = 0
  Throw_Sum = rep(0,Throw_Num)
  while (counter < Dice_Num){
    #Throw the Dice for the number of times you specify
    Throw_Dice = replicate(Throw_Num, sample(1:6,1,replace = TRUE))
    
    #Sum up the Throws
    Throw_Sum = Throw_Sum + Throw_Dice
    
    #Add the Counter
    counter = counter + 1
  }
  return(Throw_Sum)
}
```

Doing a Random Experiment and Visualize
=========================================================
- Do a Random Experiment
```{r}
# Throw the 6 Dice 60 times
Throw_Sum = Dice_Throw_Sum(60,6)

#Visualize
plot(Throw_Sum)

```

Doing a Random Experiment Several times and Plot
=========================================================
- Do a Random Experiment
```{r}
# Throw the 6 Dice 600 times
Throw_Sum = Dice_Throw_Sum(600,6)

#Visualize
plot(Throw_Sum)

```

Doing a Random Experiment and Estimate Frequency
=========================================================
- Do a Random Experiment
```{r}
# Throw the 6 Dice 6000 times
Throw_Sum = Dice_Throw_Sum(6000,6)

#Visualize
hist(Throw_Sum)
```

Doing a Random Experiment and Counting Frequency
=========================================================
- Do a Random Experiment
```{r}
# Throw the 6 Dice 60000 times
Throw_Sum = Dice_Throw_Sum(6000,6)

# Getting the position of 34,26 & 36
OP34or26or36 = which(Throw_Sum==34|Throw_Sum==26|Throw_Sum==36)

# Counting the number of cases for 34, 26 or 36
Throw34or26or36 = length(OP34or26or36)
```

Doing a Random Experiment and Getting Probabilities
=========================================================
- Calculate the Proportions from the Large Sample
```{r}
WinProb = Throw34or26or36/length(Throw_Sum)
print(WinProb)
LoseProb = 1 - WinProb
```

Calculate the Expectations from the Random Experiment
=========================================================
- Calculate the Probability*Payoff
```{r}
#Expected Payoff from 1 round of Play
Expectations = 1000*WinProb + (-60)*LoseProb

#Payoff Expectations from 10 round of Play
print(Expectations*10)
```

Run the Actual Experiment
=========================================================
- Declare the Payoff and Run the experiment
```{r}
#Declaring the Payoffs before you start the Experiment
Win_list = c(34,26,36)
Win_Payoff = 1000
Lose_Payoff = -60

#Play the Games and Get the OP sum for Throw of 6 dice
Get_Play_OP = replicate(10,Dice_Throw_Sum(1,6))
print(Get_Play_OP)
```

Compute the Payoffs
=========================================================
- Calculate the Payoff
```{r}
#Check if Payoff Exists
Payoff_Exists = which(Get_Play_OP==34|Get_Play_OP==26|Get_Play_OP==36)
WinNos = length(Payoff_Exists)
print(WinNos)

#Calculate Actual Payoff
Final_Payoff = WinNos*Win_Payoff + (10 - WinNos)*Lose_Payoff
print(Final_Payoff)

```

Using the Function "match"
============================================================================================
- The basic "match" function takes in 2 arguments - a source vector & a vector to be matched
- Wherever the Source Vector finds the matching vector, it pulls out the Location in the matching vector
- Wherever there is no match, it shows NA
- The syntax is as below  match(Source Vector, Vector to be Matched)

Alternate Payoff Deduction
============================================================================================
- Calculate the Wins and Loses
```{r}
#Check if Payoff Exists
print(Get_Play_OP)
Payoff_Exists_Match = match(Get_Play_OP,Win_list)
print(Payoff_Exists_Match)
LoseNos = sum(is.na(Payoff_Exists_Match))
print(LoseNos)
```

Calculate Payoff
=============================================================
```{r}
#Calculate Actual Payoff
Final_Payoff = LoseNos*Lose_Payoff + (10 - LoseNos)*Win_Payoff
print(Final_Payoff)
```

Calculate the Expectations for No Success
=========================================================
- Calculate the Bernouli Trial
```{r}
# Something here is definitely wrong??
NoWin = dbinom(0,10,WinProb)
print(NoWin)
Expectations = -600*NoWin + 1000*(1-NoWin)

print(Expectations)
```

Googly Time
==========================================================
- The game you played has 11 possible outcomes as given below:
- -600  (0 Wins)
- 460     1520    2580
- 3640    4700    5760
- 6820    7880    8940
- 10,000 (10 Wins)
-
- You have 10 in 11 chances of winning the game. So should you have played the game?


Calculate the Expectations through a Function
=========================================================
```{r}
Get_PLay_Expectations = function(Number_of_Trials,Winning_Probability, WinPayoff, LosePayoff)
{
  count = 0
  FinalPayoff = 0
  
  while (count <= Number_of_Trials)
  {
    # Calculating the Payoff in different scenarios of Win & Lose
    currPayoff = WinPayoff*count + LosePayoff*(Number_of_Trials - count)
    
    Winning_Prob = dbinom(count,Number_of_Trials,Winning_Probability)
    
    FinalPayoff = FinalPayoff + currPayoff*Winning_Prob
    
    count = count + 1
  }
  return(FinalPayoff)
}

```


Getting your Expectations Correct
=========================================================
```{r}
PlayExpectations = Get_PLay_Expectations(10,Winning_Probability = WinProb,WinPayoff = 1000,LosePayoff = -60)
print(PlayExpectations)

```

Compute some basic Statistics
================================================
```{r}
DiceDistribution = Dice_Throw_Sum(60000,6)
# Getting the Sum of the Distribution
sum(DiceDistribution)

# Getting the Mean of the Distribution
mean(DiceDistribution)

# Getting the Median of the Distribution
median(DiceDistribution)

```

Compute some basic Statistics
================================================
```{r}
# Getting the Maximum of the Distribution
max(DiceDistribution)

# Getting the Minimum of the Distribution
min(DiceDistribution)

# Showing the variance of the Distribution
var(DiceDistribution)

```

Compute Summary Statistics
================================================
```{r}
# Getting the first few elements of the Distribution
head(DiceDistribution)

# Showing the last few elements of the Distribution
tail(DiceDistribution)

# Getting the Summary of the Distribution
summary(DiceDistribution)

```

Compute Summary Statistics
================================================
```{r}
# Tabulating the elements distribution
table(DiceDistribution)
DiceDist = table(DiceDistribution)

```

Compute Summary Statistics
================================================
```{r}
# Tabulating the elements distributions
hist(DiceDistribution, xlab = "Dice Output", ylab = "No. of Observation")

```

Compute Summary Statistics
================================================
```{r}
# Getting the location of the maximum elements of the Distribution
which(DiceDist==max(DiceDist))
which.max(DiceDist)

# Getting the location of the minimum elements of the Distribution
which(DiceDist==min(DiceDist))
which.min(DiceDist)

```

Compute Summary Statistics
================================================
```{r}
# Getting the Element with maximum count from the Distribution
names(which.max(DiceDist))

# Getting the Element with minimum count from the Distribution
names(which.min(DiceDist))

```

Getting the Element with Max Frequency
================================================
```{r}
# Getting element with Maximum Frequency
DiceDist[which(DiceDist==max(DiceDist))]

# Getting element with Minimum Frequency
DiceDist[which(DiceDist==min(DiceDist))]

```

Getting the Top 3 Elements
================================================
```{r}
# Getting element with Minimum Frequency
head(sort(DiceDist),3)

#Getting element with Maximum Frequency
tail(sort(DiceDist),3)

```

Alt: Getting the Top 3 Elements
================================================
```{r}
# Getting element with Maximum Frequency
DiceDist[order(DiceDist, decreasing = TRUE)][1:3]

# Getting element with Minimum Frequency
DiceDist[order(DiceDist, decreasing = FALSE)][1:3]

```

Using table to check some conditions
================================================
```{r}
# Getting element with Maximum Frequency
table(DiceDistribution>26)

DiceDist[names(DiceDist)==34|names(DiceDist)==26|names(DiceDist)==36]

```

Using table to check some conditions
================================================
```{r}
DiceDist[as.numeric(DiceDist) >= 4000 & as.numeric(DiceDist) <= 5000]
```


Problem Prototype - Simplifying the Complex Function
=========================================================
- A function to simulate the summing of the Dice
```{r}
# Declare the Function
Dice_Throw_Game = function (Friend_Num, Game_Num) 
{
  Throw_Dice = replicate(Friend_Num, sample(1:6, size=Game_Num, replace = TRUE))
  
  return(Throw_Dice)
  
}

```
Problem Prototype - Simplifying the Complex Function
=========================================================
- A function to simulate the summing of the Dice
```{r}
# Call the Function

Dice_Throw_Game(Friend_Num=2,Game_Num=4)
Dice_Throw_Game(Friend_Num=4,Game_Num=2)

```

Problem Prototype - Simplifying the Complex Function
=========================================================
- A function to simulate the summing of the Dice
```{r}
# Declare the Function
Dice_Game_Sum = function (Friend_Num, Game_Num) 
{
  
  Dice_Matrix = Dice_Throw_Game(Friend_Num, Game_Num)
  
  df_Throw_Dice = data.frame(Dice_Matrix)
  df_Throw_Dice$throw_sum = rowSums(df_Throw_Dice)
  
  return(df_Throw_Dice$throw_sum)
  
}

Dice_Game_Sum(Friend_Num=2,Game_Num=4)
```

Doing a Random Experiment and Counting Frequency
=========================================================
- Do a Random Experiment
```{r}

# Throw the 6 Dice 60000 times
Throw_Sum = Dice_Game_Sum(Friend_Num=6, Game_Num=10)

print (Throw_Sum)

```

Getting the True Cases
=========================================================
- Do a Random Experiment
```{r}
WinList = c(34,26,36)

# Getting the presence of 34,26 & 36
OP34or26or36 = Throw_Sum %in% WinList

print(OP34or26or36)

```


Doing a Random Experiment and Getting Probabilities
=========================================================
- Calculate the Proportions from the Large Sample
```{r}
WinProb = sum(OP34or26or36)/length(OP34or26or36)
print(WinProb)
LoseProb = 1 - WinProb
```

Calculate the Expectations from the Random Experiment
=========================================================
- Calculate the Probability*Payoff
```{r}
#Expected Payoff from 1 round of Play
Expectations = 1000*WinProb + (-60)*LoseProb

#Payoff Expectations from 10 round of Play
print(Expectations*10)
```

Run the Actual Experiment
=========================================================
- Declare the Payoff and Run the experiment
```{r}
#Declaring the Payoffs before you start the Experiment
Win_list = c(34,26,36)
Win_Payoff = 1000
Lose_Payoff = -60

#Play the Games and Get the OP sum for Throw of 6 dice
Get_Play_OP = Dice_Game_Sum(6,6)
print(Get_Play_OP)
```




The Expand-Grid function
============================================================================================
- Used to expand vectors inside the function
- Prepares all possible combination of the vectors


The Expand-Grid Example
=============================================================================================
```{r}
tossin = c("Head","Tail")
chillr = c("Re 1", "Rs 2")

# Getting the expansion from the coin toss
expand.grid(face=tossin)

# Getting the expansion from the chiller
expand.grid(coin=chillr)
```

The Expand-Grid Example
============================================================================================
```{r}
toss = c("Head","Tail")
chillr = c("Re 1", "Rs 2")

# Getting the expansion from the coin toss
expand.grid(face=tossin, coin=chillr)
```

The Expand-Grid Example
=============================================================================================
```{r}
rSeq=c(10:12)

# Getting the expansion from the coin toss
expand.grid(face=tossin, coin=chillr, mySeq=rSeq)
```


Using Expand-Grid
========================================================
Six of you Navya, Bhavya, Kabya, Amit, Sumit & Punit are playing the game of throwing a dice. How many total combinations are there. Find the total possible cases which gives you the output as 21

Expand Grid
=================================================
```{r}
pDie = c(1:6)
myDiceThrow = expand.grid(Navya = pDie,Bhavya = pDie, Kabya = pDie, Amit = pDie, Punit = pDie, Sumit = pDie)

#Finding the Class of the Output Container
class(myDiceThrow)

#Finding the Dimensions
dim(myDiceThrow)
```

Expand Grid
=================================================
```{r}

#Finding the header rows
names(myDiceThrow)

#Checking how the output looks like
tail(myDiceThrow)
```

Alt-Commands for Dimensions
=================================================
```{r}
# Find the number of Columns
ncol(myDiceThrow)

# Find the number of Rows
nrow(myDiceThrow)

```

R Expand Grid
=================================================
```{r}
# Preparing the condition where they got the sum as 21
myJackpotBin = subset(myDiceThrow, (Navya+Bhavya+Kabya+Amit+Punit+Sumit==21))

# Finding the total cases supporting the sum as 21
nrow(myJackpotBin)

head(myJackpotBin)

```

Getting the condition as a Data Frame column
=================================================
```{r}
# Preparing the condition where they got the sum as 21

myDiceThrow$throw_sum = rowSums(myDiceThrow)
  
myJackpotBin = subset(myDiceThrow, throw_sum==21)

# Finding the total cases supporting the sum as 21
nrow(myJackpotBin)

```

