Analytics Tools & Techniques for Decision Making
========================================================
author: Prof Sourav Saha
date: 
width: 1280
height: 800


Principles of Data Science with R
========================================================
- Understand the Problem
- Create outline for the Solution
- Capture Data
- Visualize Data
- Build viable prototype
- Run the experiment
- Evaluate your prototype
- Extend to the entire solution
- Interpret

The basics of R
========================================================
- Creating a dataset
- Basic operation on the dataset
- Mean, Standard Deviation, Correlation
- Plotting


Illustrations 
========================================================
## Here we will use R to show basic operations

```{r}

# Get the Working Directory
getwd()

# Define a list for Age
age = c(1,3,5,2,11,9,3,9,12,3)

# Define a list for Weight
weight = c(4.4,5.3,7.2,5.2,8.5,7.3,6.0,10.4,10.2,6.1)


```

Illustrations 
========================================================
## Here we will use R to show basic operations

```{r}

# Find the mean Weight
mean(weight)

# Find the Standard Deviation
sd(weight)

```

Illustrations 
========================================================
## Here we will use R to show basic operations

```{r}
# Define a summary for Weight
summary(weight)

```


Illustrations 
========================================================
## Here we will use R to show basic operations

```{r}

# Draw histogram for Weight
hist(weight)

```

Bivariate insights 
========================================================
## Here we will use R to get basic information

```{r}

# Find the Correlation
cor(age,weight)

```

Bivariate Plots 
========================================================
## Here we will use R to get basic information

```{r}

# Plot the Data
plot(age,weight)

```

Data Frames
========================================================
## Here we will use R to get basic information

```{r}

new_df = data.frame(age,weight)

str(new_df)

```

Working on Plots
========================================================
## Here we will use R to get basic information

```{r}

dose  = c(20, 30, 40, 45, 60)
drugA = c(16, 20, 27, 40, 60)
drugB = c(15, 18, 25, 31, 40)

plot(dose, drugA, type="l")
# Type is line

```


Working on Plots
========================================================
## Here we will use R to get basic information

```{r}

plot(dose, drugB, type="l", pch=23, lty=6, col="blue")
# Given the pixel size, line-type, color and background

```



Working on Dataframes
========================================================
## Here we will use R to get basic information

```{r}

dose  = c(20, 30, 40, 45, 60)
drugA = c(16, 20, 27, 40, 60)
drugB = c(15, 18, 25, 31, 40)

dose_df=data.frame(dose,drugA,drugB)

```

Working on Dataframes
========================================================
## Here we will use R to get basic information

```{r}

plot(dose_df$dose, dose_df$drugA, type="b", 
     main="Clinical Trials for Drug A", 
     xlab="Dosage", ylab="Drug Response",
     xlim=c(0, 60), ylim=c(0, 70)
     )

```

Pairwise Scatter Plot
========================================================
## Drawing a Scatter Plot pairwise

```{r}

# Scatter-plot matrices
pairs(~ dose + drugA + drugB, data=dose_df, 
      main="Basic Scatterplot Matrix")
```

Working on Dataframes
========================================================
## Here we will use R to get basic information

```{r}

opar = par(no.readonly=TRUE)
par(mfrow=c(2,2))
plot(dose_df$dose,dose_df$drugA, main="Scatterplot of Dose vs vs. Drug A")
plot(dose_df$dose,dose_df$drugB, main="Scatterplot of Dose vs vs. Drug B")

hist(dose_df$dose, main="Histogram of Dosages")
boxplot(dose_df$dose, main="Boxplot of Dosages")
par(opar)

```

Calculations using R
========================================================
- Operator 	Description
- addition (+)
- subtraction (-)
- multiplication (*)
- division (/)
- exponentiation (^ or **)
- remainder x %% y 	modulus (x mod y) 5%%2 is 1
- quotient x %/% y 	integer division 5%/%2 is 2



R Calculator 
========================================================
## Here we will use R to Add, Subtract and Multiply

```{r}

# Addition
1+2

# Subtraction
1-2

# Multiplication
2*3

```

R Calculator 
========================================================
## Here we will use R to Divide

```{r}

# Division
6/2

6/4

# Remainder operation
6%%4

# Quotient operation
6%/%4

# Exponentiation
2^3


```

Creating Variables in R
========================================================
## Assigning variables

```{r}
vADD = 1+2

v_SUB = 1-2

v_MUL = 2*3

Div1 = 6/2

Div2 = 6/4
```


Rules for Creating Variables in R
========================================================
- Cannot start with a number
- Except underscore(_) and dot(.), no other special characters permitted
- Can contain number(s) in the middle or the end


Printing numerical values in R
========================================================
## Printing the numerical values

```{r}

print(1+2*2)

vOP = 1+2*2
print(vOP)

v1REM = 6%%4
print(v1REM)

```


Looking at the R print statement
========================================================
```{r}

# Printing a statement
print("Happy to learn R")

# Printing a statement using Variables
cStatement = "I am actively learning new concepts in R"
print(cStatement)

```
